# Quiz 3

## Question 1
In a population of interest, a sample of 9 men yielded a sample average brain volume of 1,100cc and a standard deviation of 30cc. What is a 95% Students T confidence interval for the mean brain volume in this new population?

* [1031, 1169]
* [1092, 1108]
* **[1077,1123]**
* [1080, 1120]

### Answer
#### T Confidence intervals

- Creating a confidence interval using the CLT takes the form $Est \pm ZQ \times SE_{Est}$
- Gosset's $t$ distribution and $t$ confidence intervals are for small samples
- They are of the form $Est \pm TQ \times SE_{Est}$
- These are some of the handiest of intervals
- If you want a rule between whether to use a $t$ interval
or normal interval, just always use the $t$ interval

#### Gosset's $t$ distribution

- Invented by William Gosset (under the pseudonym "Student") in 1908
- Has thicker tails than the normal
- Is indexed by a degrees of freedom; gets more like a standard normal as df (degrees of freedom) gets larger
- It assumes that the underlying data are iid 
Gaussian with the result that
$$
\frac{\bar X - \mu}{S/\sqrt{n}}
$$
follows Gosset's $t$ distribution with $n-1$ degrees of freedom
- (If we replaced $s$ by $\sigma$ the statistic would be exactly standard normal)
- Interval is $\bar X \pm t_{n-1} S/\sqrt{n}$ where $t_{n-1}$
is the relevant quantile

```{r}
n <- 9
s <- 30
x <- 1100

# calculate quantile, 2.5% on each side
p <- .95 # (0.975) is 95% with 2.5% on both sides of the range
quantile <- p+(1-p)/2

ci = x + c(-1, 1) * qt(quantile, df=n-1) * s / sqrt(n)
ci
```

## Question 2
A diet pill is given to 9 subjects over six weeks. The average difference in weight (follow up - baseline) is -2 pounds. What would the standard deviation of the difference in weight have to be for the upper endpoint of the 95% T confidence interval to touch 0?

* **2.60**
* 0.30
* 1.50
* 2.10

### Answer
Interval \(CI_{up} = \bar X + t_{n-1} S / \sqrt{n}\), where \(t_{n-1}\) is the relevant quantile

Rewritten, to get standard deviation: \(S = \frac{CI_{up} - \bar X * \sqrt{n}}{t_{n-1}}\)

```{r}
n <- 9
diff <- -2
ci <- 0
# calculate quantile, 2.5% on each side
p <- .95 # (0.975) is 95% with 2.5% on both sides of the range
quantile <- p+(1-p)/2

s <- (ci - diff * sqrt(n)) / qt(quantile, df = n-1)
round(s, 2)
```

## Question 3
In an effort to improve running performance, 5 runners were either given a protein supplement or placebo. Then, after a suitable washout period, they were given the opposite treatment. Their mile times were recorded under both the treatment and placebo, yielding 10 measurements with 2 per subject. The researchers intend to use a T test and interval to investigate the treatment. Should they use a paired or independent group T test and interval?

* Independent groups, since all subjects were seen under both systems
* **A paired interval**
* You could use either
* It's necessary to use both

### Answer
Each test subject have results for each treatment.

## Question 4
In a study of emergency room waiting times, investigators consider a new and the standard triage systems. To test the systems, administrators selected 20 nights and randomly assigned the new triage system to be used on 10 nights and the standard system on the remaining 10 nights. They calculated the nightly median waiting time (MWT) to see a physician. The average MWT for the new system was 3 hours with a variance of 0.60 while the average MWT for the old system was 5 hours with a variance of 0.68. Consider the 95% confidence interval estimate for the differences of the mean MWT associated with the new system. Assume a constant variance. What is the interval? Subtract in this order (New System - Old System).

* [1.29, 2.70]
* [1.25, 2.75]
* [-2,70, -1.29]
* **[-2.75, -1.25]**

### Answer

Independent group T intervals

- Therefore a $(1 - \alpha)\times 100\%$ confidence interval for $\mu_y - \mu_x$ is 
$$
    \bar Y - \bar X \pm t_{n_x + n_y - 2, 1 - \alpha/2}S_p\left(\frac{1}{n_x} + \frac{1}{n_y}\right)^{1/2}
$$
- The pooled variance estimator is $$S_p^2 = \{(n_x - 1) S_x^2 + (n_y - 1) S_y^2\}/(n_x + n_y - 2)$$ 
- Rewritten for available variables: $$S_p = \sqrt{\{(n_x - 1) var_x + (n_y - 1) var_y\}/(n_x + n_y - 2)}$$
- Remember this interval is assuming a constant variance across the two groups
- If there is some doubt, assume a different variance per group, which we will discuss later

```{r}
n_new <- 10     # size of population for new system
n_old <- 10
v_new <- 0.6    # variance for new system
v_old <- 0.68
x_new <- 3      # median waiting time for new system
x_old <- 5

# calculate quantile, 2.5% on each side
p <- .95 # (0.975) is 95% with 2.5% on both sides of the range
quantile <- p+(1-p)/2

# calculate pooled standard deviation
s <- sqrt(((n_new - 1) * v_new + (n_old - 1) * v_old)/(n_new + n_old - 2))

ci <- (x_new-x_old) + c(-1, 1) * qt(quantile, df=n_new+n_old-2) * s * (1/n_new + 1/n_old)^0.5
round(ci,2)

ci <- (x_new-x_old) + c(-1, 1) * qt(quantile, df=n_new+n_old-2) * s * sqrt(1/n_new + 1/n_old)
round(ci,2)
```

## Question 5
Suppose that you create a 95% T confidence interval. You then create a 90% interval using the same data. What can be said about the 90% interval with respect to the 95% interval?

* It is impossible to tell.
* The interval will be wider
* **The interval will be narrower.**
* The interval will be the same width, but shifted.

### Answer
Increased certainty results in a wider range, less certainty = narrower range

## Question 6
To further test the hospital triage system, administrators selected 200 nights and randomly assigned a new triage system to be used on 100 nights and a standard system on the remaining 100 nights. They calculated the nightly median waiting time (MWT) to see a physician. The average MWT for the new system was 4 hours with a standard deviation of 0.5 hours while the average MWT for the old system was 6 hours with a standard deviation of 2 hours. Consider the hypothesis of a decrease in the mean MWT associated with the new treatment. What does the 95% independent group confidence interval with unequal variances suggest vis a vis this hypothesis? (Because there's so many observations per group, just use the Z quantile instead of the T.)

* When subtracting (old - new) the interval is entirely above zero. The new system does not appear to be effective.
* When subtracting (old - new) the interval contains 0. There is not evidence suggesting that the new system is effective.
* **When subtracting (old - new) the interval is entirely above zero. The new system appears to be effective.**
* When subtracting (old - new) the interval contains 0. The new system appears to be effective.

### Answer
```{r}
n_new <- 100    # size of population for new system
n_old <- 100
s_new <- 0.5    # standard deviation for new system
s_old <- 2
x_new <- 4      # median waiting time for new system
x_old <- 6

# calculate quantile, 2.5% on each side
p <- .95 # (0.975) is 95% with 2.5% on both sides of the range
quantile <- p+(1-p)/2

# calculate the variance from the standard deviation (so i can use the same code from Q4, just need to switch the x_new-x_old around and use qnorm instead of qt)
v_new <- s_new^2
v_old <- s_old^2

# calculate pooled standard deviation
s <- sqrt(((n_new - 1) * v_new + (n_old - 1) * v_old)/(n_new + n_old - 2))

# the question asked about old-new
ci <- (x_old-x_new) + c(-1, 1) * qnorm(quantile) * s * (1/n_new + 1/n_old)^0.5
round(ci,2)

# also have a look at new-old
ci <- (x_new-x_old) + c(-1, 1) * qnorm(quantile) * s * (1/n_new + 1/n_old)^0.5
round(ci,2)
```

## Question 7
Suppose that 18 obese subjects were randomized, 9 each, to a new diet pill and a placebo. Subjects body mass indices (BMIs) were measured at a baseline and again after having received the treatment or placebo for four weeks. The average difference from follow-up to the baseline (followup - baseline) was -3 kg/m2 for the treated group and 1 kg/m2 for the placebo group. The corresponding standard deviations of the differences was 1.5 kg/m2 for the treatment group and 1.8 kg/m2 for the placebo group. Does the change in BMI over the four week period appear to differ between the treated and placebo groups? Assuming normality of the underlying data and a common population variance, calculate the relevant 90% t confidence interval. Subtract in the order of (Treated - Placebo) with the smaller (more negative) number first.

* [2.469, 5.531]
* [-5.531, -2.469]
* [2.636, 5.364]
* **[-5.364, -2.636]**

### Answer
```{r}
n_treated <- 9    # subjects taking drug
n_placebo <- 9
s_treated <- 1.5  # kg/m2 standard deviation
s_placebo <- 1.8
x_treated <- -3   # kg/m2 average BMI change
x_placebo <- 1

# calculate quantile, 2.5% on each side
p <- .90 # 5% on both sides of the range
quantile <- p+(1-p)/2

# calculate the variance from the standard deviation
v_treated <- s_treated^2
v_placebo <- s_placebo^2

# calculate pooled standard deviation
s <- sqrt(((n_treated - 1) * v_treated + (n_placebo - 1) * v_placebo)/(n_treated + n_placebo - 2))

ci <- (x_treated-x_placebo) + c(-1, 1) * qt(quantile, df=n_placebo+n_treated-2) * s * (1/n_treated + 1/n_placebo)^0.5
round(ci,3)
```
